Selecting Computations: Theory and Applications

### 2012-06-26 paper/vago.uai/paper.tex
current state

\newcommand {\term}[1] {\textbf{#1}}
\newcommand {\note}[1] {\textcolor{red}{[[#1]]}}
\newcommand {\Evidence} {\mathcal{E}}
\newcommand {\R} {\mathbb{R}}
\newcommand {\maps} {\colon}
\newcommand {\given} {\mathrel{|}}
\newcommand {\sample} {\sim}
\newcommand {\iid} {\stackrel{\rm iid}{\sample}}
\DeclareMathOperator*{\argmax}{argmax}

@@

@@

\author{}
%\author{ {\bf Nicholas Hay} and {\bf Stuart Russell} \\   
%Computer Science Division \\  
%University of California\\ 
%Berkeley, CA 94720 \\ 
%\And 
%{\bf Solomon Eyal Shimony} and {\bf David Tolpin}  \\ 
%Department of Computer Science \\ 
%Ben-Gurion University of the Negev\\
%Beer Sheva, Israel
%}

@@

\begin{abstract}
Sequential decision problems are often approximately solvable by
simulating possible future action sequences.  {\em Metalevel} decision
procedures have been developed for selecting {\em which} action
sequences to simulate, based on some estimate of the expected
improvement in decision quality resulting from any particular
simulation; an example is the recent work on using selection rules
derived for bandit problems to control Monte Carlo tree search in the
game of Go.  In this paper we develop a theoretical basis for
metalevel decisions in the statistical framework of {\em selection problems}, 
arguing that this is more appropriate than the bandit
framework.  We derive a number of basic results applicable to simple
Monte Carlo selection problems, including the first finite sampling
bounds for optimal policies in certain cases; we also provide a simple
counterexample to the intuitive conjecture that an optimal policy will
necessarily reach a decision in all cases.  We then derive heuristic
approximations in both Bayesian and distribution-free settings and
demonstrate their superiority to bandit-based heuristics in one-shot
decision problems and in Go.
\end{abstract}

@@

\section{Upper bounds on value of information}\label{approx-nonbayesian-section}

@@

Theorem \ref{thm:optimal-myopic} provides necessary conditions
for a stopping condition for the optimal policy in terms of a myopic
policy, but that does not preclude premature stopping of a greedy policy.
Conversely, Theorem \ref{thm:myopic-optimal} provides sufficient conditions, but these 
conditions are hard to evaluate in practice.

The myopic selection policy uses
the intrinsic value of information (VOI) $\Lambda_i$ of testing an arm $i$, which is
the expected reward of the best arm given the information obtained by the test,
minus the expected reward of the best arm according to the current information state.
Unfortunately, the pure \textit{myopic} VOI estimate is of little use in
Monte-Carlo sampling, since the effect of a single sample is small,
and the myopic VOI estimate will often be zero.

Instead, we generalize myopic policies to include ``semi-myopic"
up to $N$ samples of the same arm; these are called ``blinkered" policies.
Such policies allow potentially unlimited lookahead, but only in a single ``direction'' (one specific arm),
as if we ``had our blinkers on''.

@@

In many practical applications of the selection problem, such as search in
the game of Go, prior distributions are unavailable\footnote{The analysis is also applicable to
some Bayesian settings, using ``fake" samples to simulate prior distributions.}
In such cases, one can still bound
the value of information w.r.t. myopic policies using {\em concentration
inequalities} to derive distribution-independent bounds on the
VOI. We obtain such bounds under the
following assumptions:
\begin{enumerate}
\item Samples are iid given the value of the arms (variables), as in the Bayesian schemes such as Bernoulli
sampling.
\item The expectation of a selection in a belief state is equal to the sample mean (and therefore,
   after sampling terminates, the arm with the greatest sample mean will be selected).
\end{enumerate}

When considering possible samples in the ``blinkered'' semi-myopic setting,
two cases are possible:
\begin{itemize}
\item the arm $\alpha$ with the highest sample mean $\overline
  X_\alpha$ is tested, and $\overline X_\alpha$ becomes lower than
  $\overline X_\beta$ of the second-best arm $\beta$;
\item another arm $i$ is tested, and $\overline X_i$ becomes higher
than $\overline X_\alpha$.
\end{itemize}

Our bounds below are applicable to any bounded distribution (with rewards being
bounded to $[0,1]$). Similar
bounds can be derived for certain unbounded distributions, such as the
normally distributed prior of the value with normally distributed
sampling.
We derive a VOI bound for testing an arm $N$ times,
where $N$ can be the remaining ``budget'' of available samples or
any other integer quantity.
Denote by  $\Lambda_i^b$ the intrinsic VOI of testing the $i$th arm
$N$ times, and the number of
samples already taken from the $i$th arm by $n_i$.
\begin{thm} $\Lambda_i^b$ is bounded from above as
\begin{align}
\label{eqn:thm-be}
  \Lambda_\alpha^b&\le \frac {N \overline X_\beta^{n_\beta}} {n_\alpha} \Pr(\overline X_\alpha^{n_\alpha+N}\le\overline X_\beta^{n_\beta})\nonumber\\
\Lambda_{i|i\ne\alpha}^b&\le \frac {N(1-\overline X_\alpha^{n_\alpha})} {n_i}\Pr(\overline   X_i^{{n_i}+N}\ge\overline X_\alpha^{n_\alpha})
\end{align}
\label{thm:be}
\end{thm}
\vspace{-2em}
\begin{proof} For the case $i\ne \alpha$, the probability that the
  $i$th arm is finally chosen instead of $\alpha$ is
  $\Pr(\overline X_i^{n_i+N} \ge \overline X_\alpha^{n_\alpha})$. $X_i \le 1$,
  therefore $\overline X_i^{n_i+N}\le \overline
  X_\alpha^{n_\alpha}+\frac {N(1-\overline X_\alpha^{n_\alpha})} {N+n_i}$. Hence, the intrinsic value of blinkered
  information is at most: 
\begin{align}
\label{eq:simplistic}
\frac{ N(1-\overline  X_\alpha^{n_\alpha})}
  {N+n_i}&\Pr(\overline X_i^{{n_i}+N}\hspace{-0.5em}\ge\overline X_\alpha^{n_\alpha})\nonumber \\
&\le\frac{ N(1-\overline  X_\alpha^{n_\alpha})}
{n_i}\Pr(\overline X_i^{{n_i}+N}\hspace{-0.5em}\ge\overline X_\alpha^{n_\alpha})
\end{align}
  Proof for the case $i=\alpha$ is similar.
\end{proof}

The probabilities can be bounded from above using the
Hoeffding inequality \citep{Hoeffding.ineq}:
\begin{thm} The probabilities in equations (\ref{eqn:thm-be}) are bounded from above as
\begin{align}
  \label{eqn:probound-blnk-hoeffding}
  \Pr&(\overline X_\alpha^{{n_\alpha}+N} \le \overline X_\beta^{n_\beta})
  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} - \overline X_\beta^{n_\beta})^2 n_\alpha
  \right)\nonumber\\
  \Pr&(\overline X_{i|i\ne\alpha}^{n_\alpha+N} \ge \overline X_\beta^{n_\beta})
  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} -\overline  X_i^{n_i})^2 n_i \right)
\end{align}
where $\varphi=\min \left(2(\frac {1+n/N} {1+\sqrt {n/N}})^2\right)=8(\sqrt 2 - 1)^2 > 1.37$.
\label{thm:hoeffding-prob-bounds}
\end{thm}

\begin{comment}
	\begin{proof}
	Equations (\ref{eqn:probound-blnk-hoeffding}) follow from the
	observation that if $i\ne\alpha$, $\overline X_i^{n_i+N}>\overline X_\alpha^{n_i}$
	if and only if the mean $\overline X_i^N$ of $N$ samples from $n_i+1$
	to $n_i+N$ is at least $\overline X_\alpha^{n_i}+(\overline X_\alpha^{n_i}-\overline
	X_i^{n_i})\frac {n_i} N$.

	For any $\delta$, the probability that $\overline X_i^{n_i+N}$ is greater
	than $\overline X_\alpha^{n_i}$ is less than the probability that
	$\IE[X_i]\ge\overline X_i^n+\delta$
	\emph{or} $\overline X_i^N\ge \IE[X_i]+\overline X_\alpha^{n_\alpha}
	- \overline X_i^{n_i} - \delta +(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})\frac {n_i} N$,
	thus, by the union bound, less than the sum of the probabilities:
	\begin{align}
	\Pr&(\overline X_i^{n_i+N}\ge \overline X_\alpha^{n_i})\nonumber\\
	   &\le\Pr(\IE[X_i]-\overline X_i^{n_i} \ge \delta)\\
	   &\hspace{0.25em}+\Pr\left(\overline X_i^N\hspace{-0.5em} - \IE[X_i] \ge \overline X_\alpha^{n_\alpha}\hspace{-0.5em}
	           - \overline X_i^{n_i}\hspace{-0.5em} - \delta +(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})\frac {n_i} N\right)\nonumber
	\end{align}
	Bounding the probabilities on the right-hand side using the Hoeffding
	inequality, obtain:
	\begin{align}
	\Pr&(\overline X_i^{n_i+N}\ge \overline X_\alpha^{n_\alpha})\le \nonumber\\
	    &\exp(-2\delta^2n_i)+\nonumber\\
	    &\quad\exp\left(-2\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
	         - \overline X_i^{n_i})\left(1+\frac {n_i} N\right)
	         - \delta\right)^2N\right)
	\label{eqn:app-hoeffding-le-maxexp}
	\end{align}
	Find $\delta$ for which the two terms on the right-hand side of
	(\ref{eqn:app-hoeffding-le-maxexp}) are equal:
	\begin{equation}
	\exp(-\delta^2n) = \exp\left(-2\left((\overline X_\alpha - \overline X_i)(1+\frac {n_i} N) - \delta\right)^2N\right)\label{eqn:app-hoeffding-eq-exp}
	\end{equation}
	Solve (\ref{eqn:app-hoeffding-eq-exp}) for $\delta$: $\delta=\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
	- \overline X_i^{n_i}) \ge 2(\sqrt 2 - 1)(\overline X_\alpha^{n_\alpha}\hspace{-0.5em}-\overline X_i^{n_i})$. Substitute $\delta$ into 
	(\ref{eqn:app-hoeffding-le-maxexp}) and obtain
	\begin{align}
	\Pr&(\overline X_i^{n_i}\ge \overline X_\alpha^{n_\alpha}) \nonumber\\
	& \le 2\exp\left(-2\left( \frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}
	                          (\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})\right)^2 n_i\right)\nonumber \\
	& \le 2\exp(-8(\sqrt 2 - 1)^2(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})^2n_i)\nonumber\\
	& = 2\exp(-\varphi(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})^2n_i)
	\end{align}
	Derivation for the case $i=\alpha$ is similar.
	\end{proof}	
\end{comment}

\begin{crl}
An upper bound on the VOI estimate $\Lambda_i^b$ is obtained
by substituting (\ref{eqn:probound-blnk-hoeffding}) into (\ref{eqn:thm-be}).
\begin{align}
  \label{eqn:bound-blnk-hoeffding}
  \Lambda&_\alpha^b \le \hat\Lambda_\alpha^b=\frac {2N\overline X_\beta^{n_\beta}} {n_\alpha}\exp\left(- \varphi(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_\beta^{n_\beta})^2 n_\alpha\right)\nonumber\\
  \Lambda&_{i|i\ne\alpha}^b\le \hat\Lambda_i^b=  \frac {2N(1-\overline  X_\alpha^{n_\alpha})} {n_i}\exp\left(- \varphi(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})^2 n_i\right)
\end{align}
\label{crl:bound-blnk-hoeffding}
\end{crl}
\vspace{-2em}

More refined bounds can be obtained through tighter estimates on the
probabilities in Equation \ref{eq:simplistic}, for example, based on the empirical Bernstein
inequality~\citep{MaurerPontil.benrstein}, or through a more careful
application of the Hoeffding inequality, resulting in:
\begin{align}
\Lambda_i^b&\le\frac {N\sqrt \pi} {n_i \sqrt {n_i}}
  \left[\mathrm{erf}\left((1\hspace{-0.25em}-\hspace{-0.25em}\overline X_i^{n_i})\sqrt {n_i}\right)
      \hspace{-0.25em}-\hspace{-0.25em}\mathrm{erf}\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})\sqrt{n_i}\right)\right]\nonumber\\
\Lambda_\alpha^b&\le\frac {N\sqrt \pi} {n_\alpha \sqrt {n_\alpha}}
  \left[\mathrm{erf}\left(\overline X_\alpha^{n_\alpha}\sqrt {n_\alpha}\right)
      -\mathrm{erf}\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_\beta^{n_\beta})\sqrt{n_\alpha}\right)\right]
\label{eqn:erf-blinkered}
\end{align}

\section{Practical sampling control: empirical evaluation}
\label{sec:empirical-evaluation}\label{mcts-section}

As the selection MDP is intractable, we use myopic or semi-myopic policies instead.
The myopic policies usually separate out the decision of {\em whether} to
sample or to stop (called the stopping policy), and {\em what} to sample.
We first examine the latter issue.

\subsection{Selecting The Best Arm}
\label{sec:emp-arm}

Assuming that the sample costs are constant,
a semi-myopic policy will decide to test the arm that has the best
current VOI estimate. 
When the distributions are unknown, it makes sense
to use the upper bounds established in Theorem~\ref{thm:be} instead,
which is the method adopted here.
This evaluation assumes a fixed budget of samples, which is
completely used up by each of the candidate schemes, making a stopping
criterion irrelevant.

\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{flat.pdf}
\caption{Random instances: regret vs. number of samples}
\label{fig:random-instances}
\end{figure}

The sampling policies are compared on random Multi-armed bandit 
selection problem instances. Figure~\ref{fig:random-instances} shows results for
randomly-generated Multi-armed bandits with 25 Bernoulli arms, with
the mean rewards of the arms distributed uniformly in the range~$[0,
  1]$, for a range of sample budgets~$200..2000$, with multiplicative
step of~$2$. The experiment for each number of samples was repeated
10000 times. UCB1, as well as policies  based on bounds
(\ref{eqn:bound-blnk-hoeffding}) (VOI) and
(\ref{eqn:erf-blinkered}) (VOI+) are compared. UCB1 is always considerably worse than the
VOI-aware sampling policies.

\subsection{Sampling in trees}

Sampling in trees is considerably more complicated.
The goal of Monte-Carlo tree search \citep{Chaslot.montecarlo} at the root node 
is usually to select an action that appears to be the best based on outcomes
of \textit{search rollouts}.
But the goal of rollouts at non-root nodes
is different than at the root: here it is important to better approximate the ``true'' 
value of the node, so that selection at the root can be more informed. The exact analysis
of sampling at internal nodes is outside the scope of this paper. At present we 
have no better proposal for internal nodes than to use UCT there.

We thus propose the following hybrid sampling scheme:
\begin{itemize}
\item at the root node, sample based on the VOI estimate;
\item at non-root nodes, sample using UCT.
\end{itemize}

Strictly speaking, even at the root node the assumptions underlying our belief-state
MDP for selection do not hold exactly. UCT is an adaptive scheme, and therefore the values
generated by sampling at non-root nodes will typically cause valued observed at
children of the root node to be non-stationary. 
Nevertheless, sampling based on VOI estimates
computed as for stationary distributions work well in
practice. As illustrated
by the empirical evaluation (Section~\ref{sec:empirical-evaluation}),
estimates based on upper bounds on the VOI result in good sampling
policies, which exhibit performance comparable to the performance of
some state-of-the-art heuristic algorithms.

\subsubsection{Stopping criterion}
\label{sec:control-stopping-criterion}

When a sample has a known cost commensurable with the value of
information of a measurement, an upper bound on the intrinsic VOI can also
be used to stop the sampling if the intrinsic VOI of any action
is less than the cost of sampling $C$:
\begin{equation}
\mbox{stop if } \max_i \Lambda_i \le C
\end{equation}
VOI estimates~(\ref{eqn:thm-be},
\ref{eqn:bound-blnk-hoeffding}) include the remaining sample budget $N$ as a
factor, but given the cost of a single sample $c$, the cost of the
remaining samples accounted for in estimating the intrinsic VOI is
$C=cN$. $N$ can be dropped on both sides of the inequality,
and a reasonable stopping criterion would be:
\begin{align}
\frac 1 N \Lambda_\alpha^b \le&\frac {\overline X_\beta^{n_\beta}}
  {n_\alpha}\Pr(\overline X_\alpha^{n_\alpha+N}\le\overline
  X_\beta^{n_\alpha})\le c\nonumber\\
\frac 1 N \max_i\Lambda_i^b\le &\max_i\frac {(1-\overline X_\alpha^{n_\alpha})} {n_i}\Pr(\overline
  X_i^{n_i+N}\ge\overline X_\alpha^{n_\alpha})\le c\nonumber\\
    &\forall i: i\ne\alpha
\label{eqn:stopping-blnk}
\end{align}
The empirical evaluation (Section~\ref{sec:empirical-evaluation})
confirms the viability of this stopping criterion and illustrates the
influence of the sample cost $c$ on the performance of
the sampling policy. When the sample cost $c$ is unknown, one can perform initial calibration experiments
to determine a reasonable value, as done in this paper.

\subsubsection{Sample redistribution in trees}
\label{sec:control-redistribution}

>>>>

\newcommand {\term}[1] {\textbf{#1}}
\newcommand {\note}[1] {\textcolor{red}{[[#1]]}}
\newcommand {\Evidence} {\mathcal{E}}
\newcommand {\R} {\mathbb{R}}
\newcommand {\maps} {\colon}
\newcommand {\given} {\mid} %{\mathrel{|}}
\newcommand {\sample} {\sim}
\newcommand {\iid} {\stackrel{\rm iid}{\sample}}
\newcommand {\Nat} {\mathbb{N}}
\DeclareMathOperator*{\argmax}{argmax}

@@

\excludecomment{hiddenproof}

@@

%\author{}
\author{ {\bf Nicholas Hay} and {\bf Stuart Russell} \\   
Computer Science Division \\  
University of California\\ 
Berkeley, CA 94720 \\ 
\And 
{\bf Solomon Eyal Shimony} and {\bf David Tolpin}  \\ 
Department of Computer Science \\ 
Ben-Gurion University of the Negev\\
Beer Sheva, Israel
}

@@

\input{abstract}

@@

\section{Upper bounds on Value of Information}\label{approx-nonbayesian-section}

@@

\begin{comment}
	Theorem \ref{thm:optimal-myopic} provides necessary conditions
	for a stopping condition for the optimal policy in terms of a myopic
	policy, but that does not preclude premature stopping of a greedy policy.
	Conversely, Theorem \ref{thm:myopic-optimal} provides sufficient conditions, but these 
	conditions are hard to evaluate in practice.

	The myopic selection policy uses
	the intrinsic value of information (VOI) $\Lambda_i$ of testing an arm~$i$, which is
	the expected reward of the best arm given the information obtained by the test,
	minus the expected reward of the best arm according to the current information state.
	Unfortunately, the pure \textit{myopic} VOI estimate is of little use in
	Monte-Carlo sampling, since the effect of a single sample is small,
	and the myopic VOI estimate will often be zero.

	Instead, we generalize myopic policies to include ``semi-myopic"
	up to $N$ samples of the same arm; these are called ``blinkered" policies.
	Such policies allow potentially unlimited lookahead, but only in a single ``direction'' (one specific arm),
	as if we ``had our blinkers on''.	
\end{comment}

@@

In many practical applications of the selection problem, such as search in
the game of Go, prior distributions are unavailable.\footnote{The analysis is also applicable to
some Bayesian settings, using ``fake" samples to simulate prior distributions.}
In such cases, one can still bound
the value of information of myopic policies using {\em concentration
inequalities} to derive distribution-independent bounds on the
VOI. We obtain such bounds under the
following assumptions:
\begin{enumerate}
\item Samples are iid given the value of the arms (variables), as in the Bayesian schemes such as Bernoulli
sampling.
\item The expectation of a selection in a belief state is equal to the sample mean (and therefore,
   after sampling terminates, the arm with the greatest sample mean will be selected).
\end{enumerate}

When considering possible samples in the blinkered semi-myopic setting,
two cases are possible: either
	the arm $\alpha$ with the highest sample mean $\overline
  	X_\alpha$ is tested, and $\overline X_\alpha$ becomes lower than
 	$\overline X_\beta$ of the second-best arm $\beta$;
or, 
	another arm~$i$ is tested, and $\overline X_i$ becomes higher
    than $\overline X_\alpha$.

Our bounds below are applicable to any bounded distribution (without loss of generality 
bounded in $[0,1]$). Similar
bounds can be derived for certain unbounded distributions, such as the
normally distributed prior value with normally distributed
sampling.
We derive a VOI bound for testing an arm a fixed $N$ times,
where $N$ can be the remaining budget of available samples or
any other integer quantity.
Denote by  $\Lambda_i^b$ the intrinsic VOI of testing the $i$th arm
$N$ times, and the number of
samples already taken from the $i$th arm by $n_i$.
\begin{thm} $\Lambda_i^b$ is bounded from above as
\begin{align}
\label{eqn:thm-be}
  \Lambda_\alpha^b&\le \frac {N \overline X_\beta^{n_\beta}} {n_\alpha} \Pr(\overline X_\alpha^{n_\alpha+N}\le\overline X_\beta^{n_\beta})\nonumber\\
\Lambda_{i|i\ne\alpha}^b&\le \frac {N(1-\overline X_\alpha^{n_\alpha})} {n_i}\Pr(\overline   X_i^{{n_i}+N}\ge\overline X_\alpha^{n_\alpha})
\end{align}
\label{thm:be}
\end{thm}
\begin{hiddenproof}
	\vspace{-2em}
	\begin{proof} For the case $i\ne \alpha$, the probability that the
	  $i$th arm is finally chosen instead of $\alpha$ is
	  $\Pr(\overline X_i^{n_i+N} \ge \overline X_\alpha^{n_\alpha})$. $X_i \le 1$,
	  therefore $\overline X_i^{n_i+N}\le \overline
	  X_\alpha^{n_\alpha}+\frac {N(1-\overline X_\alpha^{n_\alpha})} {N+n_i}$. Hence, the intrinsic value of blinkered
	  information is at most: 
	\begin{align}
	\label{eq:simplistic}
	\frac{ N(1-\overline  X_\alpha^{n_\alpha})}
	  {N+n_i}&\Pr(\overline X_i^{{n_i}+N}\hspace{-0.5em}\ge\overline X_\alpha^{n_\alpha})\nonumber \\
	&\le\frac{ N(1-\overline  X_\alpha^{n_\alpha})}
	{n_i}\Pr(\overline X_i^{{n_i}+N}\hspace{-0.5em}\ge\overline X_\alpha^{n_\alpha})
	\end{align}
	  Proof for the case $i=\alpha$ is similar.
	\end{proof}		
\end{hiddenproof}

The probabilities can be bounded from above using the
Hoeffding inequality \citep{Hoeffding.ineq}:
\begin{thm} The probabilities in \eqref{eqn:thm-be} are bounded from above as
\begin{align}
  \label{eqn:probound-blnk-hoeffding}
  \Pr&(\overline X_\alpha^{{n_\alpha}+N} \le \overline X_\beta^{n_\beta})
  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} - \overline X_\beta^{n_\beta})^2 n_\alpha
  \right)\nonumber\\
  \Pr&(\overline X_{i|i\ne\alpha}^{n_\alpha+N} \ge \overline X_\beta^{n_\beta})
  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} -\overline  X_i^{n_i})^2 n_i \right)
\end{align}
where $\varphi=\min \left(2(\frac {1+n/N} {1+\sqrt {n/N}})^2\right)=8(\sqrt 2 - 1)^2 > 1.37$.
\label{thm:hoeffding-prob-bounds}
\end{thm}

\begin{hiddenproof}
	\begin{proof}
	\eqref{eqn:probound-blnk-hoeffding}) follow from the
	observation that if $i\ne\alpha$, $\overline X_i^{n_i+N}>\overline X_\alpha^{n_i}$
	if and only if the mean $\overline X_i^N$ of $N$ samples from $n_i+1$
	to $n_i+N$ is at least $\overline X_\alpha^{n_i}+(\overline X_\alpha^{n_i}-\overline
	X_i^{n_i})\frac {n_i} N$.

	For any $\delta$, the probability that $\overline X_i^{n_i+N}$ is greater
	than $\overline X_\alpha^{n_i}$ is less than the probability that
	$\IE[X_i]\ge\overline X_i^n+\delta$
	\emph{or} $\overline X_i^N\ge \IE[X_i]+\overline X_\alpha^{n_\alpha}
	- \overline X_i^{n_i} - \delta +(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})\frac {n_i} N$,
	thus, by the union bound, less than the sum of the probabilities:
	\begin{align}
	\Pr&(\overline X_i^{n_i+N}\ge \overline X_\alpha^{n_i})\nonumber\\
	   &\le\Pr(\IE[X_i]-\overline X_i^{n_i} \ge \delta)\\
	   &\hspace{0.25em}+\Pr\left(\overline X_i^N\hspace{-0.5em} - \IE[X_i] \ge \overline X_\alpha^{n_\alpha}\hspace{-0.5em}
	           - \overline X_i^{n_i}\hspace{-0.5em} - \delta +(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})\frac {n_i} N\right)\nonumber
	\end{align}
	Bounding the probabilities on the right-hand side using the Hoeffding
	inequality, obtain:
	\begin{align}
	\Pr&(\overline X_i^{n_i+N}\ge \overline X_\alpha^{n_\alpha})\le \nonumber\\
	    &\exp(-2\delta^2n_i)+\nonumber\\
	    &\quad\exp\left(-2\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
	         - \overline X_i^{n_i})\left(1+\frac {n_i} N\right)
	         - \delta\right)^2N\right)
	\label{eqn:app-hoeffding-le-maxexp}
	\end{align}
	Find $\delta$ for which the two terms on the right-hand side of
	\eqref{eqn:app-hoeffding-le-maxexp} are equal:
	\begin{equation}
	\exp(-\delta^2n) = \exp\left(-2\left((\overline X_\alpha - \overline X_i)(1+\frac {n_i} N) - \delta\right)^2N\right)\label{eqn:app-hoeffding-eq-exp}
	\end{equation}
	Solve \eqref{eqn:app-hoeffding-eq-exp} for $\delta$: $\delta=\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
	- \overline X_i^{n_i}) \ge 2(\sqrt 2 - 1)(\overline X_\alpha^{n_\alpha}\hspace{-0.5em}-\overline X_i^{n_i})$. Substitute $\delta$ into 
	\eqref{eqn:app-hoeffding-le-maxexp} and obtain
	\begin{align}
	\Pr&(\overline X_i^{n_i}\ge \overline X_\alpha^{n_\alpha}) \nonumber\\
	& \le 2\exp\left(-2\left( \frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}
	                          (\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})\right)^2 n_i\right)\nonumber \\
	& \le 2\exp(-8(\sqrt 2 - 1)^2(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})^2n_i)\nonumber\\
	& = 2\exp(-\varphi(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})^2n_i)
	\end{align}
	Derivation for the case $i=\alpha$ is similar.
	\end{proof}	
\end{hiddenproof}

\begin{crl}
An upper bound on the VOI estimate $\Lambda_i^b$ is obtained
by substituting \eqref{eqn:probound-blnk-hoeffding} into \eqref{eqn:thm-be}.
\begin{align}
  \label{eqn:bound-blnk-hoeffding}
  \Lambda&_\alpha^b \le \hat\Lambda_\alpha^b=\frac {2N\overline X_\beta^{n_\beta}} {n_\alpha}\exp\left(- \varphi(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_\beta^{n_\beta})^2 n_\alpha\right)\nonumber\\
  \Lambda&_{i|i\ne\alpha}^b\le \hat\Lambda_i^b=  \frac {2N(1-\overline  X_\alpha^{n_\alpha})} {n_i}\exp\left(- \varphi(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})^2 n_i\right)
\end{align}
\label{crl:bound-blnk-hoeffding}
\end{crl}
\vspace{-2em}

More refined bounds can be obtained through tighter estimates on the
probabilities in \eqref{eqn:thm-be}, for example, based on the empirical Bernstein
inequality~\citep{MaurerPontil.benrstein}, or through a more careful
application of the Hoeffding inequality, resulting in:
\begin{align}
\Lambda_i^b&\le\frac {N\sqrt \pi} {n_i \sqrt {n_i}}
  \left[\mathrm{erf}\left((1\hspace{-0.25em}-\hspace{-0.25em}\overline X_i^{n_i})\sqrt {n_i}\right)
      \hspace{-0.25em}-\hspace{-0.25em}\mathrm{erf}\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})\sqrt{n_i}\right)\right]\nonumber\\
\Lambda_\alpha^b&\le\frac {N\sqrt \pi} {n_\alpha \sqrt {n_\alpha}}
  \left[\mathrm{erf}\left(\overline X_\alpha^{n_\alpha}\sqrt {n_\alpha}\right)
      -\mathrm{erf}\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_\beta^{n_\beta})\sqrt{n_\alpha}\right)\right]
\label{eqn:erf-blinkered}
\end{align}

Selection problems usually separate out the decision of {\em whether} to
sample or to stop (called the stopping policy), and {\em what} to sample.
We'll examine the first issue here, along with the empirical evaluation 
of the above approximate algorithms, and the second in the following section.

Assuming that the sample costs are constant,
a semi-myopic policy will decide to test the arm that has the best
current VOI estimate. 
When the distributions are unknown, it makes sense
to use the upper bounds established in \thmref{thm:be}, as we do in the following.
This evaluation assumes a fixed budget of samples, which is
completely used up by each of the candidate schemes, making a stopping
criterion irrelevant.

\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{flat.pdf}
\caption{Average regret of various policies as a function of the fixed number 
of samples in a 25-action Bernoulli sampling problem, over 10000 trials.}
\label{fig:random-instances}
\end{figure}

The sampling policies are compared on random Bernoulli
selection problem instances. \figref{fig:random-instances} shows results for
randomly-generated selection problems with 25 Bernoulli arms, where
the mean rewards of the arms are distributed uniformly in~$[0,1]$, 
for a range of sample budgets~$200..2000$, with multiplicative
step of~$2$, averaging over 10000 trials.  We compare UCB1 with the 
policies based on the bounds in
\eqref{eqn:bound-blnk-hoeffding} (VOI) and
\eqref{eqn:erf-blinkered} (VOI+).
UCB1 is always considerably worse than the VOI-aware sampling policies.

\section{Sampling in trees}
\label{sec:empirical-evaluation}\label{mcts-section}

The previous section addressed the selection problem in the flat case.
Selection in trees is more complicated.  The goal of Monte-Carlo tree 
search \citep{Chaslot.montecarlo} at the root node 
is usually to select an action that appears to be the best based on outcomes
of \textit{search rollouts}.
But the goal of rollouts at non-root nodes
is different than at the root: here it is important to better approximate the
value of the node, so that selection at the root can be more informed. The exact analysis
of sampling at internal nodes is outside the scope of this paper. At present we 
have no better proposal for internal nodes than to use UCT there.

We thus propose the following hybrid sampling scheme: 
	at the \emph{root node}, sample based on the VOI estimate;
	at \emph{non-root nodes}, sample using UCT.

Strictly speaking, even at the root node the stationarity assumptions\footnote{This is not a restriction,
however, of the general formalism in \secref{sec:optimal}.} 
underlying our belief-state
MDP for selection do not hold exactly. UCT is an adaptive scheme, and therefore the values
generated by sampling at non-root nodes will typically cause values observed at
children of the root node to be non-stationary. 
Nevertheless, sampling based on VOI estimates
computed as for stationary distributions works well in
practice. As illustrated
by the empirical evaluation (\secref{sec:empirical-evaluation}),
estimates based on upper bounds on the VOI result in good sampling
policies, which exhibit performance comparable to the performance of
some state-of-the-art heuristic algorithms.

\subsection{Stopping criterion}
\label{sec:control-stopping-criterion}

When a sample has a known cost commensurable with the value of
information of a measurement, an upper bound on the intrinsic VOI can also
be used to stop the sampling if the intrinsic VOI of any arm
is less than the total cost of sampling $C$:
\begin{equation}
\mbox{stop if } \max_i \Lambda_i \le C
\end{equation}
The VOI estimates of \eqrefs{eqn:thm-be}{eqn:bound-blnk-hoeffding} 
include the remaining sample budget $N$ as a
factor, but given the cost of a single sample $c$, the cost of the
remaining samples accounted for in estimating the intrinsic VOI is
$C=cN$. $N$ can be dropped on both sides of the inequality,
giving a reasonable stopping criterion:
\begin{align}
\frac 1 N \Lambda_\alpha^b \le&\frac {\overline X_\beta^{n_\beta}}
  {n_\alpha}\Pr(\overline X_\alpha^{n_\alpha+N}\le\overline
  X_\beta^{n_\alpha})\le c\nonumber\\
\frac 1 N \max_i\Lambda_i^b\le &\max_i\frac {(1-\overline X_\alpha^{n_\alpha})} {n_i}\Pr(\overline
  X_i^{n_i+N}\ge\overline X_\alpha^{n_\alpha})\le c\nonumber\\
    &\forall i: i\ne\alpha
\label{eqn:stopping-blnk}
\end{align}
The empirical evaluation (\secref{sec:empirical-evaluation})
confirms the viability of this stopping criterion and illustrates the
influence of the sample cost $c$ on the performance of
the sampling policy. When the sample cost $c$ is unknown, one can perform initial calibration experiments
to determine a reasonable value, as done in the following.

\subsection{Sample redistribution in trees}
\label{sec:control-redistribution}

The above hybrid approach assumes
that the information obtained from rollouts in the
current state is discarded after an real-world action is selected. In practice,
many successful Monte-Carlo tree search algorithms reuse rollouts
generated at earlier search states, if the sample traverses the
current search state during the rollout; thus, the value of information of a rollout is
determined not just by the influence on the choice of the action at
the current state, but also by its potential influence on the choice at future
search states.
